Deciding Bitvector Termination with Program Synthesis

### 2014-07-02 papers/termination/paper.tex
Whitespace

\newcommand{\xmark}{\ding{55}}
\newcommand{\todo}[1]{{\bf TODO:} #1}
\newcommand{\newC}{C$^-$\xspace}

@@

 The main contributions of our work can be summarised as follows:

\begin{itemize}
\item We designed a technique for computing ranking functions that correctly accounts for the wrap-around behavior caused by under- and overflows in bit-vector and floating point arithmetic. To the best of our knowledge, this is the first approach able to compute ranking functions for programs handling floats. Our technique is not restricted to finding linear ranking functions, but can also compute (lexicographic) non-linear  ones. We justified the need for such a non restrictive procedure by computing the probability ... .
\item  We rephrased the termination problem as a second-order satisfaction problem and made 
use of results in genetic programming to efficiently solve it. We have also investigated the effects of genetic operators on the search space for ranking functions and computed theoretical 
bounds on the convergence time ...
\item We present a formulation that allows us to prove termination without an expensive reachability check (as we don't use Ramsey-based termination arguments).  In particular,
we only need a bounded model checker that performs a single unwinding of a loop.
%% \item Our technique is able to uniformly handle conditionally-terminating loops, as well as programs with
%% multiple loops.

@@

\item bit level termination and non-termination analysis. This enables
  the precise treatment of bit-wise operations and arithmetic modulo
  fixed widths.

@@

Given that our state space is finite, a transition relation induces a infinite execution if and only if some state is 
visited twice in the course of the execution. Hence, an alternative formulation of the non-termination problem 
for bit-vectors may search for a state that is visited twice.
However, this would require reasoning about the transitive closure of the transition relation.\\

>>>>

\newcommand{\xmark}{\ding{55}}
\newcommand{\todo}[1]{{\bf TODO:} #1}

@@

 The main contributions of our work can be summarised as follows:
\begin{itemize}
\item We designed a technique for computing ranking functions that correctly accounts for the wrap-around behavior caused by under- and overflows in bit-vector and floating point arithmetic. To the best of our knowledge, this is the first approach able to compute ranking functions for programs handling floats. Our technique is not restricted to finding linear ranking functions, but can also compute (lexicographic) non-linear  ones. We justified the need for such a non restrictive procedure by computing the probability ... .
\item  We rephrased the termination problem as a second-order satisfaction problem and made 
use of results in genetic programming to efficiently solve it. We have also investigated the effects of genetic operators on the search space for ranking functions and computed theoretical 
bounds on the convergence time ...
\item We present a formulation that allows us to prove termination without an expensive reachability check (as we don't use Ramsey-based termination arguments).  In particular,
we only need a bounded model checker that performs a single unwinding of a loop.
%% \item Our technique is able to uniformly handle conditionally-terminating loops, as well as programs with
%% multiple loops.

@@

@@

Given that our state space is finite, a transition relation induces a infinite execution if and only if some state is 
visited infinitely many times in the course of the execution. Hence, an alternative formulation of the non-termination problem 
for bit-vectors may search for singleton recurrence sets (the recurrence set contains only the state visited infinitely many times).
However, this would require reasoning about the transitive closure of the transition relation.\\
